---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```


[![R build status](https://github.com/MarselScheer/bootGOF/workflows/R-CMD-check/badge.svg)](https://github.com/MarselScheer/bootGOF/actions)
[![Coverage Status](https://img.shields.io/codecov/c/github/MarselScheer/bootGOF/develop.svg)](https://codecov.io/github/MarselScheer/bootGOF?branch=develop)
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html)
[![Project Status: Active – The project has reached a stable, usable
state and is being actively
developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/bootGOF)](https://cran.r-project.org/package=bootGOF)
[![metacran downloads](https://cranlogs.r-pkg.org/badges/bootGOF)](https://cran.r-project.org/package=bootGOF)
[![license](https://img.shields.io/badge/license-GPL--3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0.en.html)


# bootGOF

Bootstrap based goodness-of-fit tests for (linear) models. 
Assume you have fitted a statistical model, e.g. classical linear
model or generalized linear model or a model that follows
$Y = m(\beta^\top X) + \epsilon$. This package allows to perform
a rigorous statistical test to check if the chosen model family is correct.
    
## Example 

First we generate a data-set in order to apply the package. 

```{r}
set.seed(1)
N <- 100
X1 <- rnorm(N)
X2 <- rnorm(N)
d <- data.frame(
  y = rpois(n = N, lambda = exp(4 + X1 * 2 + X2 * 6)),
  x1 = X1,
  x2 = X2)
```
Note that both covariates influence the dependent variable $Y$. Taking
only one of the covariates into account obviously leads to a model family
that is not correct and the GOF-test should reveal that:

```{r}
fit <- glm(y ~ x1, data = d, family = poisson())

library(bootGOF)
mt <- GOF_model(
  model = fit,
  data = d,
  nmb_boot_samples = 100,
  simulator_type = "parametric",
  y_name = "y",
  Rn1_statistic = Rn1_KS$new())
mt$get_pvalue()
```

On the other hand assuming the correct model family should in general not
be rejected by the GOF-test:

```{r}
fit <- glm(y ~ x1 + x2, data = d, family = poisson())
mt <- GOF_model(
  model = fit,
  data = d,
  nmb_boot_samples = 100,
  simulator_type = "parametric",
  y_name = "y",
  Rn1_statistic = Rn1_KS$new())
mt$get_pvalue()
```

...

## Installation

You can install it from CRAN

```{r, eval = FALSE}
install.packages("bootGOF")
```

or github

```{r gh-installation, eval = FALSE}
devtools::install_github("MarselScheer/bootGOF")
```

## Tests

After installing the package you can execute the unit tests of the
package in your environment by calling:

```{r, eval = FALSE}
library(tinytest)
tinytest::test_package("bootGOF")
```

# sessionInfo

```{r}
sessionInfo()
```
